---
title: "HW2"
author: "Megan Tran"
date: '`October 3, 2022'
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install-libraries}
library(tidytext)
library(tidyverse)
library(dplyr)
library(dtplyr)
library(ggplot2)
library(data.table)
```

#Read in the data
```{r read-data, cache=TRUE}
if (!file.exists("individual.csv")) {
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_individual.csv", "individual.csv", method="libcurl", timeout = 60) 
}
ind <- data.table::fread("individual.csv")

str(ind)
```

```{r read-data-2, cache=TRUE}
if (!file.exists("regional.csv")) {
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_regional.csv", "regional.csv", method="libcurl", timeout = 60) 
}
reg <- data.table::fread("regional.csv")

str(reg)
```


#Merging the 2 datasets, making sure there's no duplicates
```{r}
nrow(reg)
nrow(ind)
```


```{r merging-data}
data <-
merge(
  # Data
  x     = ind,      
  y     = reg, 
  # List of variables to match
  by.x  = "townname",
  by.y  = "townname", 
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
  ) 
```

##Data Wrangling

#Question 1 Impute data
```{r}
nrow(data)
```

There are still 1200 rows in the merged dataset.

#missing values? 
```{r}
mean(is.na(data))
```
Although the proportion of missing values is low, I'll be imputing it with the average from "male" "hispanic".

#impute missing values with average from "male" and "hispanic"
```{r impute-bmi}
data[, bmi_imp := fcoalesce(bmi, mean(bmi, na.rm = TRUE)),
by = .(male,hispanic)]
```

```{r impute-fev}
data[, fev_imp := fcoalesce(fev, mean(fev, na.rm = TRUE)),
by = .(male,hispanic)]
```

```{r impute-asthma}
data[, asthma_imp := fcoalesce(as.numeric(asthma), mean(asthma, na.rm = TRUE)),
by = .(male,hispanic)]
```


#Question 2 
Create a new categorical variable named “obesity_level” using the BMI measurement (underweight BMI<14; normal BMI 14-22; overweight BMI 22-24; obese BMI>24). To make sure the variable is rightly coded, create a summary table that contains the minimum BMI, maximum BMI, and the total number of observations per category.

```{r}
data$obesity_level <- as.factor(ifelse(data$bmi_imp<14, 'Underweight',
                              ifelse(data$bmi_imp>=14 & data$bmi_imp<22, 'Normal',
                              ifelse(data$bmi_imp>=22 & data$bmi_imp<=24, 'Overweight',
                              ifelse(data$bmi_imp>24, 'Obese','Missing')))))
```

```{r}
data[, .(
    min_bmi      = min(bmi_imp, na.rm=TRUE),
    max_bmi   = max(bmi_imp, na.rm=TRUE), count = length(bmi_imp)
    ),
    by = obesity_level
    ][order(obesity_level)] 
```

#Question 3 
Create another categorical variable named “smoke_gas_exposure” that summarizes “Second Hand Smoke” and “Gas Stove.” The variable should have four categories in total.

```{r}
data$smoke_gas_exposure <- as.factor(ifelse(data$smoke==1 & data$gasstove==1, 'Both',
                                     ifelse(data$smoke==1 & data$gasstove==0, 'SmokeOnly',
                                            ifelse(data$smoke==0 & data$gasstove==1, 'GasOnly',
                                                   ifelse(data$smoke==0 & data$gasstove==0, 'None', 'Missing')))))
```

```{r}
summary(data$smoke_gas_exposure)
```

#Question 4
Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”

```{r}
data[, .(
    mean_fev     = mean(fev_imp, na.rm=TRUE),
    sd_fev       = sd(fev_imp, na.rm=TRUE), 
    prop_asthma  = mean(asthma_imp, na.rm=TRUE),
    sd_asthma    = sd(asthma_imp, na.rm=TRUE)
    ),
    by = townname
    ][order(townname)] 
```
```{r}
data[, .(
    mean_fev     = mean(fev_imp, na.rm=TRUE),
    sd_fev       = sd(fev_imp, na.rm=TRUE), 
    prop_asthma  = mean(asthma_imp, na.rm=TRUE),
    sd_asthma    = sd(asthma_imp, na.rm=TRUE)
    ),
    by = male
    ]
```
```{r}
data[, .(
    mean_fev     = mean(fev_imp, na.rm=TRUE),
    sd_fev       = sd(fev_imp, na.rm=TRUE), 
    prop_asthma  = mean(asthma_imp, na.rm=TRUE),
    sd_asthma    = sd(asthma_imp, na.rm=TRUE)
    ),
    by = obesity_level
    ][order(obesity_level)] 
```

```{r}
data[, .(
    mean_fev     = mean(fev_imp, na.rm=TRUE),
    sd_fev       = sd(fev_imp, na.rm=TRUE), 
    prop_asthma  = mean(asthma_imp, na.rm=TRUE),
    sd_asthma    = sd(asthma_imp, na.rm=TRUE)
    ),
    by = smoke_gas_exposure
    ][order(smoke_gas_exposure)] 
```


##Looking at the Data (EDA)

EDA Checklist

Formulate a question: 1. What is the association between BMI and FEV (forced expiratory volume)? 2. What is the association between smoke and gas exposure and FEV? 3. What is the association between PM2.5 exposure and FEV?
Read in the data: done
Check the dimensions and headers and footers of the data: 

```{r}
dim(data)
head(data)
tail(data)
```

Check the variable types in the data: done

Take a closer look at some/all of the variables: done
Validate with an external source
Conduct some summary statistics to answer the initial question

#Question 1.
Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.

```{r}
ggplot(data = data, mapping = aes(x = bmi_imp, y = fev_imp, color=townname)) + 
  geom_point() + 
  geom_smooth(method = lm, col = "black") +
  labs(title = "Scatterplots of BMI vs FEV by Town") + 
  labs(x = expression("Body Mass Index"), y = "Forced Expiratory Volume") +
  facet_wrap(~ townname, nrow = 3, ncol=4)
```
Based on these scatterplots, it can be gathered that BMI and FEV are positively correlated for those in all 12 towns.

#Question 2 
Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.

```{r}
ggplot(data=data) +
  geom_histogram(mapping = aes(x = fev_imp, fill = obesity_level)) +
        scale_fill_brewer(palette = "Pastel1")
```  
There are a significantly greater amount of those with "Normal" BMI than any other category. The second largest group our "Obese". But there isn't a clear association between obesity level and fev.

```{r}
ggplot(data=data) +
  geom_histogram(mapping = aes(x = fev_imp, fill = smoke_gas_exposure)) + 
        scale_fill_brewer(palette = "Set2") 
```
It seems like second hand smoke and gas exposure together result in the highest FEV. Being exposed to gas only is closely second. Therefore, there might be a correlation between FEV and smoke/gas exposure. 

#Question 3 
Barchart of BMI by smoke/gas exposure.

```{r}
ggplot(data=data) +
  geom_bar(mapping = aes(x = obesity_level, fill = smoke_gas_exposure))
```
From this barchart, I can see that a majority of people in all groups are exposed to gas only. Again, there doesn't seem to be correlation between BMI and smoke/gas exposure.



#Question 4
Statistical summary graphs of FEV by BMI (obesity level) and FEV by smoke/gas exposure category.

```{r}
data[!is.na(fev_imp)] %>%
  ggplot(mapping = aes(x=obesity_level, y=fev_imp)) + 
    stat_summary(fun.data = mean_sdl,
  geom = "errorbar")      
```
This visualization gives me the most information. I can see that as the BMI increaseses, so do the bars. Perhaps there's a correlation between the variables after all.


```{r}
data[!is.na(fev_imp)] %>%
  ggplot(mapping = aes(x=smoke_gas_exposure, y=fev_imp)) + 
    stat_summary(fun.data = mean_sdl,
  geom = "errorbar")      
```
The lowest bar is the "Both" category and shortly after is the "SmokeOnly". This graph also indicates that there might be a correlation between the variables.


#Question 5
A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.

```{r}
library(leaflet)
pm25.pal <- colorFactor(topo.colors(12), domain = data$pm25_mass)
leaflet(data) %>% 
   addTiles() %>%
   addCircles (lat = ~lat, lng = ~lon, color = ~pm25.pal(pm25_mass),
                label = ~paste0(pm25_mass),
                opacity=0.01, fillOpacity = 0.05, radius = 5000) %>%
addLegend('bottomleft', pal= pm25.pal, values = (data$pm25_mass),
              title ='PM 2.5 Mass', opacity=1)
```

As a general trend, the communities that are more north have smaller concentrations of PM 2.5 than the southern communities.


#Question 6 
Choose a visualization to examine whether PM2.5 mass is associated with FEV.

```{r}
ggplot(data=data) + 
  geom_point(mapping = aes(x = pm25_mass, y = fev_imp))+
  geom_smooth(mapping = aes(x = pm25_mass, y = fev_imp))
```
I don't see a correlation because the regression line is almost horizontal.


