---
title: "Lab 03"
author: "Megan Tran"
date: '`September 7, 2022'
output:
  html_document:
    df_print: paged
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Step 1 Read in the data
First download and then read in with data.table:fread()
```{r}
if (!file.exists("met_all.gz")) {
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60) 
}
met <- data.table::fread("met_all.gz") 
```

##Step 2 Check the dimensions, headers, footers. How many columns, rows are there?
```{r}
dim(met)
## [1] 2377343      30
head(met)
tail(met)
```

## Step 3 Take a look at the variables.
```{r}
str(met)
```

## Step 4 Take a closer look at the key variables.
```{r}
table(met$year)
table(met$day)
table(met$hour)
```


```{r}
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)
```

```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```
`r max(met$elev, na.rm=TRUE)` meters

```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```
##Step #5 Check the data against an external data source.

```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```

##Step #6 Calculate summary statistics
```{r}
elev <- met[elev==max(elev)]
summary(elev)
```

```{r}
cor(elev$temp, elev$wind.sp, use="complete")
```

```{r}
cor(elev$temp, elev$hour, use="complete")
```
```{r}
cor(elev$wind.sp, elev$day, use="complete")
```
```{r}
cor(elev$wind.sp, elev$hour, use="complete")
```

```{r}
cor(elev$temp, elev$day, use="complete")
```

## Step 7. Exploratory Graphs

Look at key variables 

```{r}
hist(met$elev, breaks=100)
```
```{r}
hist(met$temp)
```

```{r}
hist(met$wind.sp) 
```

```{r}
if (!require(leaflet)) {
  install.packages("leaflet")
  library(leaflet)
}
```



```{r}
#library(leaflet)
elev <- met[elev==max(elev)]
```


```{r}
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```

```{r}
library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
```
```{r}
elev <- elev[order(date)]
head(elev)
```
```{r}
plot(elev$date, elev$temp, type='l')
```

##In August, the temperature rose and fell every 10 days. 

```{r}
plot(elev$date, elev$wind.sp, type='l')
```

## The wind trends rose and fell every 5 days towards the end of August. 

