---
title: "HW2"
author: "Megan Tran"
date: 'November 4, 2022'
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rvest)
library(httr)
library(tidyverse)
library(stringr)
library(xml2)
library (dplyr)
library(dtplyr)
library(forcats)
library(ggplot2)
library(tidytext)
```

#APIs
Using the NCBI API, look for papers that show up under the term “sars-cov-2 trial vaccine.” Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

```{r}
# Downloading the website
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")
# Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/div[1]")
# Turning it into text
counts <- as.character(counts)
# Extracting the data using regex
stringr::str_extract(counts, "[0-9,]+")
```

There are 4005 papers that have the term “sars-cov-2 trial vaccine.”

Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.

```{r}
library(httr)
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db      = "pubmed",
    term    = "sars-cov-2 trial vaccine",
    retmax  = 250
  ), 
)
# Extracting the content of the response of GET
ids <- httr::content(query_ids)

ids
```


```{r} 
# Turn the result into a character vector
ids <- as.character(ids)
# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[[:digit:]]+</Id>")[[1]]
# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- stringr::str_remove_all(ids, "</?Id>")

head(ids)
```


As we did in lab 7. Create a dataset containing the following:


```{r}
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list(
    db  = "pubmed",
    id  = paste(ids,collapse = ","),
  retmax = 250,
  rettype = "abstract"
  ))
# character vector
publications <- httr::content(publications)
publications_text <- as.character(publications)
```


Pubmed ID number:
```{r}
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
```


Title of the paper:
```{r}
titles <- str_extract(pub_char_list, "<ArticleTitle>[[:print:][:space:]]+</ArticleTitle>")
titles <- str_remove_all(titles, "</?[[:alnum:]- =\"]+>")
```


Name of the journal where it was published:
```{r}
journal <- str_extract(pub_char_list, "<Title>[[:print:][:space:]]+</Title>")
journal <- str_remove_all(journal, "[\n]")
journal <- str_remove_all(journal, "</?[[:alnum:]- =\"]+>")
```


Publication date:
```{r}
date <- str_extract(pub_char_list, "<PubDate>[[:print:][:space:]]+</PubDate>")
date <- str_remove_all(date, "</?[[:alnum:]- =\"]+>")
date <- str_remove_all(date, "[:space:]")
date <- str_remove_all(date, "[\n]")
```

Abstract of the paper (if any):
```{r}
abstract <- str_extract(pub_char_list, "<Abstract>[[:print:][:space:]]+</Abtract>")
abstract <- str_remove_all(abstract, "</?[[:alnum:]- =\"]+>")
abstract <- str_replace_all(abstract, "[[:space:]]+", " ")
```



```{r}
database <- data.frame(
         PubMedId = ids,
          Title   = titles,
          Journal = journal,
  PublicationDate = date,
         Abstract = abstract
)
knitr::kable(database, caption = "Papers about sars-cov-2 trial vaccine")
```


#Text Mining
A new dataset has been added to the data science data repository https://github.com/USCbiostats/data-science-data/tree/master/03_pubmed. The dataset contains 3241 abstracts from articles across 5 search terms. Your job is to analyse these abstracts to find interesting insights.

1. Tokenize the abstracts and count the number of each token. Do you see anything interesting? Does removing stop words change what tokens appear as the most frequent? What are the 5 most common tokens for each search term after removing stopwords?

```{r}
if (!file.exists("pubmed.csv")){
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/03_pubmed/pubmed.csv", "pubmed.csv", method="libcurl", timeout = 60)
}

data <- read.csv("pubmed.csv")
str(data)
```


```{r}
mts %>%
  unnest_tokens(word, transcription) %>%
  #anti_join(stop_words, by = c("word")) %>%
  count(word, sort = TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```


